\documentclass[a4paper,oneside,12pt]{book}
%----------------------------------------------------------------------------------------
%	WELCOME!
%   It's probably worth having a read through this file to set up the broad parameters.
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	COVER PAGE
%   The cover page is laid out in title/title.tex. You can choose a colour
%   or black and white logo
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%   Put title, author name, degree, type of work, school, department in here
%   It will be used for the title page and for the embedded PDF information
%----------------------------------------------------------------------------------------

\newcommand{\thesistitle}{Learning for sensor-based, real-time fall detection for cyclists.  } % Your thesis title, this is used in the title and abstract
\newcommand{\degree}{B. A (Mod.) Computer Science} % Your degree name, this is used in the title page and abstract
\newcommand{\typeofthesis}{Final Year Project} % dissertation, Final Year Project, report, etc.
\newcommand{\authorname}{Aidan Mongan} % Your name, this is used in the title page and PDF stuff
%% Comment out the next line if you don't want your ID to appear
\newcommand{\authorid}{14334583} % Your ID
\newcommand{\keywords}{this, that, more} % Keywords for your thesis
\newcommand{\school}{\href{http://www.scss.tcd.ie}{School of Computer Science and Statistics}} % Your school's name and URL, this is used in the title page

%% Comment out the next line if you don't want a department to appear
%\newcommand{\department}{\href{http://researchgroup.university.com}{Department Name}} % Your research group's name and URL, this is used in the title page

\AtBeginDocument{
\hypersetup{pdftitle=\thesistitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywords} % Set the PDF's keywords to your keywords
\hypersetup{pdfsubject=\degree} % Set the PDF's keywords to your keywords
}

%% Language and font encodings
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

%% Bibliographical stuff
\usepackage[round,sort,comma,numbers]{natbib}

%% Document size
% include showframe as an option if you want to see the boxes
\usepackage[a4paper,top=2.54cm,bottom=2.54cm,left=2.54cm,right=2.54cm,headheight=16pt]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\usepackage[pdftex]{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{caption} % if no caption, no colon
\usepackage{sfmath} %use sans-serif in the maths sections too
\usepackage[parfill]{parskip}    % Begin paragraphs with an empty line rather than an indent
\usepackage{setspace} % to permit one-and-a-half or double spacing
\usepackage{enumerate} % fancy enumerations like (i) (ii) or (a) (b) and suchlike
\usepackage{booktabs} % To thicken table lines
\usepackage{fancyhdr}

\pagestyle{plain} % Embrace simplicity!

%% The Mechanical engineers require your name and ID on the top of every page.
%% Uncomment the following block if you want your name and ID at the top of
%% (almost) every page.

%\pagestyle{fancy}
%\fancyhf{} % sets both header and footer to nothing
%\renewcommand{\headrulewidth}{0pt}
%\cfoot{\thepage}
%\ifdefined\authorid
%\chead{\it \authorname\ (\authorid)}
%\else
%\chead{\it \authorname}
%\fi
%% End of block

%% It is not a requirement of the university that the font should be sans-serif, but
%% the Mechanical engineers require it. Comment out the following line to disable it
\renewcommand{\familydefault}{\sfdefault} %use the sans-serif font as default

%% If you're not using sans-serif, consider using Palatino instead of the LaTeX standard
%\usepackage{mathpazo} % Use the Palatino font by default if you prefer it to Computer Modern

\renewcommand{\theequation}{\arabic{equation}} %% use continuous equation numbers

%% Format Chapter headings appropriately
\usepackage{titlesec}
\titleformat{\chapter}[hang] 
{\normalfont\huge\bfseries}{\thechapter}{1cm}{} 

\title{\thesistitle}
\author{\authorname}

\frontmatter
\begin{document}
\input{title/title.tex}
\pagenumbering{roman}
\section*{\Huge{Declaration}}
\vspace{1cm}
I hereby declare that this project is entirely my own work and that it has not been submitted as an exercise for a degree at this or any other university.

\vspace{1cm}
I have read and I understand the plagiarism provisions in the General Regulations of the University Calendar for the current year, found at \url{http://www.tcd.ie/calendar}.
\vspace{1cm}

I have also completed the Online Tutorial on avoiding plagiarism `Ready Steady Write', located at
\url{http://tcd-ie.libguides.com/plagiarism/ready-steady-write}.
\vspace{3cm}

Signed:~\rule{5cm}{0.3pt}\hfill Date:~\rule{5cm}{0.3pt}

\chapter*{Abstract}
Like all extreme sports, mountain biking comes with the potential for serious injury to the rider in the event of an accident. Non fatal injuries can easily become fatal, when one is alone, far from help and potentially incapacitated. A study conducted by Paracelsus Medical University recorded injury rates as high as 16.8 injuries per 1000 hours of riding, with 22 being moderate and 16 being severe, with rider error being the leading cause \cite{studyOfMTBInjuries}. An automated crash detection system has the potential to be life saving in the worst of circumstances.

Existing discipline-specific solutions e.g., for road only or for mountain use only, on both hardware (Specialized’s AnGi) and software (Strava Beacon,Garmin) have inflexible detection algorithms focusing on using thresholds for only one to two data points. For example AnGi records values from its inbuilt gyroscope and accelerometer, while Garmin’s system uses only accelerometer values. Such threshold-based solutions pose issues in terms of high false detection rates and a single threshold value is unlikely to be suitable for different users at different skill levels.
  
 This project expands on previous research done in the area of wearable fall detection devices for the elderly, focusing on the design of a software solution for real-time fall detection. Three data points are used: raw sensor data from both a tri-axial gyroscope and a tri-axial accelerometer as well as the rate of change of speed, calculated via GPS. The proposed system utilizes learning techniques to improve detection rates and over time generates a more personalized model. Based on pre-captured training data of both regular riding and crashes, data is classified using a multivariable logistic regression model in real-time to determine whether a crash has occurred. Raw sensor data is captured from the inbuilt sensors present on android smartphones.

This approach is implemented as an android application called “RideSafe” and was evaluated using a user study, comprising of X participants at local trail centres over a Y day period. Crash data was also collected, by means of intentional crashes in a controlled environment  for verification.  Results show that this system can successfully detect upwards of X crashes with a low rate of Y false positives….


\newpage
\onehalfspacing\raggedright %\raggedright turns off justification and hypenation

\section*{\Huge{Acknowledgements}}
Thanks Mum!

You should acknowledge any help that you have received (for example from technical staff), or input provided by, for example, a company.
\tableofcontents
%%\listoffigures
%%\listoftables
\newpage


\mainmatter
\input{introduction/introduction.tex}
\input{background/background.tex}
%%\input{project/project.tex}
\input{evaluation/evaluation.tex}
\input{conclusion/conclusion.tex}
\bibliographystyle{unsrtnat}
\bibliography{bibs/references}
\appendix
\renewcommand{\thechapter}{A\arabic{chapter}}
\input{appendix/appendix.tex}


\end{document}